Something about Performance, 2D Component of NEMO. (TBC)

### 2015-05-06 nemolite2d_perf.tex
Add plot of effect of domain size on OpenMP scaling.

We now turn to transformations related to parallelisation of the
nemolite2d code; the introduction of OpenMP and OpenACC directives. In
keeping with the \psykal approach, these transformations are
restricted to the PSy (middle) layer of the code.

@@

As with the serial optimisations, we consider the effect of each of
the OpenMP optimisation steps described in Section~\ref{sec_omp_steps}
for the $256^2$ domain. For this we principally use a single Intel,
Ivy Bridge socket which has 12 hardware cores and support for up to 24
threads with hyperthreading.
Figures~\ref{FIG_gnu_omp_256},~\ref{FIG_intel_omp_256}
and~\ref{FIG_cray_omp_256} show the performance of each of the
versions of the code on this system for the Gnu, Intel and Cray
compilers, respectively.

@@

The simplest OpenMP implementation (black lines) fails to scale beyond
two threads for any of the compilers. However, as soon as a single
PARALLEL region is introduced, the situation is greatly improved with
all three executables now scaling out to at least 12 threads (red
lines, square symbols). {\bf Add parallel efficiency lines to the OMP
  scaling plots to highlight that things are still not brilliant.}

@@

In order to quantify the scaling behaviour of the different versions
of nemolite2d with the different compilers/run-time environments, we
also plot the parallel efficiency in
Figures~\ref{FIG_gnu_omp_256},~\ref{FIG_intel_omp_256}
and~\ref{FIG_cray_omp_256}. We define parallel efficiency, $E(n)$, on
a number of threads, $n$, as:
\begin{equation}
E(n) = 100 * \frac{P(n)}{ n*P(1)}
\end{equation}
where $P(n)$ is the performance of the code on $n$ threads.

@@

>>>>

We now turn to transformations related to parallelisation of the
nemolite2d code; the introduction of OpenMP and OpenACC directives. In
keeping with the \psykal approach we do not modify either the
Algorithm- or Kernel-layer code.  Any code changes are restricted to
either the PSy (middle) layer or the underlying library that manages
e.g. the construction of field objects.

@@

As with the serial optimisations, we consider the effect of each of
the OpenMP optimisation steps described in Section~\ref{sec_omp_steps}
for the $256^2$ domain. For this we principally use a single Intel,
Ivy Bridge socket which has 12 hardware cores and support for up to 24
threads with hyperthreading (i.e. two threads per core).
Figures~\ref{FIG_gnu_omp_256},~\ref{FIG_intel_omp_256}
and~\ref{FIG_cray_omp_256} show the performance of each of the
versions of the code on this system for the Gnu, Intel and Cray
compilers, respectively.

In order to quantify the scaling behaviour of the different versions
of nemolite2d with the different compilers/run-time environments, we
also plot the parallel efficiency in
Figures~\ref{FIG_gnu_omp_256},~\ref{FIG_intel_omp_256}
and~\ref{FIG_cray_omp_256} (dashed lines and open symbols). We define
parallel efficiency (\%), $E(n)$, on a number of threads, $n$, as:
\[
E(n) = 100  \frac{P(n)}{ n P(1)}
\]
where $P(m)$ is the performance of the code on $m$ threads. For a
perfect, linearly-scaling code, $E(n)$ will be 100\%.

Since the space to explore consists of three different compilers, six
different domain sizes, five stages of optimisation and six different
thread counts we can only consider a slice through it in what follows.
In order to inform our choice of domain size,
Figure~\ref{FIG_omp_domain_size} shows the scaling behaviour of the
most performant, Intel-compiled version of nemolite2d. Although it is
common for production runs of NEMO to use MPI sub-domains of $20
\times 20$ ({\bf do we have/need a reference for that?}), it is clear
from Figure~\ref{FIG_omp_domain_size} that the limited quantity of
parallelism in the $32 \times 32$ domain inhibits scaling. Therefore,
to fully test the performance of our OpenMP implementations we have
chosen to examine results for the $256 \times 256$ domain and these
are shown in Figures~\ref{FIG_gnu_omp_256}, \ref{FIG_intel_omp_256}
and \ref{FIG_cray_omp_256}.

\begin{figure}
\centering
\includegraphics[width=\picwidth]{omp_scaling_problem_size}
\caption{The scaling behaviour of the most performant OpenMP-parallel
  version of \psykal nemolite2d for the full range of domain sizes
  considered on the CPU. Results are for the Intel compiler on a
  single Intel Ivy Bridge socket. The corresponding parallel
  efficiencies are shown using open symbols and dashed lines. The
  24-thread runs employed hyperthreading.}
\label{FIG_omp_domain_size}
\end{figure}

@@

The simplest OpenMP implementation (black lines, circle symbols) fails
to scale well for any of the compilers. For the Intel and Gnu
versions, parallel efficiency is already less than 50\% on just four
threads. The Cray version however does better and is about 45\%
efficient on eight threads.

With the move to a single PARALLEL region, the situation
is greatly improved with all three executables now scaling out to at
least 12 threads with $\sim70$\% efficiency (red lines, square symbols).

@@

@@

%\begin{figure}
%\centering
%\includegraphics[width=\picwidth]{omp_scaling_32_gnu}
%\caption{Performance of the OpenMP-parallel version of \psykal
%  nemolite2d for the Gnu compiler on a single Intel Ivy Bridge
%  socket. The corresponding parallel efficiencies are shown using open
%  symbols and dashed lines. The 24-thread runs employed hyperthreading
%  and the optimal OpenMP schedule is given in parentheses.}
%\label{FIG_gnu_omp_32}
%\end{figure}

%\begin{figure}
%\centering
%\includegraphics[width=\picwidth]{omp_scaling_32_intel}
%\caption{Performance of the OpenMP-parallel version of \psykal
%  nemolite2d for the Intel compiler on a single Intel Ivy Bridge
%  socket.  The corresponding parallel efficiencies are shown using
%  open symbols and dashed lines. The 24-thread runs employed
%  hyperthreading and the optimal OpenMP schedule is given in
%  parentheses.}
%\label{FIG_intel_omp_32}
%\end{figure}

%\begin{figure}
%\centering
%\includegraphics[width=\picwidth]{omp_scaling_32_cray}
%\caption{Performance of the OpenMP-parallel version of \psykal
%  nemolite2d for the Cray compiler on a single Intel Ivy Bridge
%  socket.  The corresponding parallel efficiencies are shown using
%  open symbols and dashed lines. The 24-thread runs employed
%  hyperthreading and the optimal OpenMP schedule is given in
%  parentheses.}
%\label{FIG_cray_omp_32}
%\end{figure}
